---
title: "190926_visiualization"
author: "Gavin Ko"
date: "9/26/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ggridges)
```
ggridges is a good plug in for ggplot that doesn't overwrite anything important
and give it a little extra merits.

Here we are at VIS and EDA!

## Create the weather data
```{r weather data loading}
weather_df = 
  rnoaa::meteo_pull_monitors(c("USW00094728", "USC00519397", "USS0023B17S"), #three sites
                      var = c("PRCP", "TMIN", "TMAX"), #humidity, minimumtemp, highesttemp
                      date_min = "2017-01-01",
                      date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY", 
                      USC00519397 = "Waikiki_HA",
                      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
weather_df

```

## Create a ggplot for that

Basic scatter plot is like this, which Jeff doesn't use pretty often.
```{r ggplot}
ggplot(weather_df, aes(x = tmin, y = tmax, col = name)) + geom_point()

```

Alternative wat to make this plot with pipe, and filter
```{r pipe}
weather_df %>% 
  filter(name == "CentralPark_NY" | name == "Waikiki_HA") %>% 
    ggplot(aes(x = tmin, y = tmax) ) + 
    geom_point( aes(col = name), alpha = 0.4 ) +
    geom_smooth( aes(col = name), se = F )
#alpha is to control the transparency of dots
#geom_smooth is to draw the trend of the general data
#se indicates not showing the gray bands

```

Facets
```{r pipe}
weather_df %>% 
  filter(name == "CentralPark_NY" | name == "Waikiki_HA") %>% 
    ggplot(aes(x = tmin, y = tmax) ) + 
    geom_point( aes(col = name), alpha = 0.4 ) +
    geom_smooth( aes(col = name), se = F ) + 
    facet_grid(. ~ name)

```

Analysis according to dates
```{r dates}
weather_df %>% 
  ggplot( aes(x = date, y = tmax, col = name) ) +
  geom_point(aes(size = prcp), alpha = 0.35) +
  geom_smooth(size = 2, se = F)
#size indicates the size of point, and can be assigned with factors

```

You can remove any layer of geom, pick and choose what you want to show.
```{r reduced}
weather_df %>% 
  ggplot( aes(x = date, y = tmax, col = name) ) +
  geom_smooth(size = 2, se = F)

```

2D scatter plot
```{r 2D}
weather_df %>% 
  ggplot( aes(x = tmin, y = tmax, col = name) ) + 
  geom_bin2d() 
#or geom_hex(), but you gotta install another package for this
```

##More kinds of plots!

```{r histogram}
weather_df %>% 
  ggplot(aes(x = tmax, fill = name)) + 
  geom_histogram(position = "dodge")
#when you indicate col >> hollow, fill >> filled color
#position dodge is to make the histograms dodge eachother

```

```{r histogram with gp comparison}
weather_df %>% 
  ggplot(aes(x = tmax, fill = name)) + 
  geom_histogram() + 
  facet_grid(. ~ name)

```

We can also put density plot, which shows a better distribution trend

```{r density}
weather_df %>% 
  ggplot(aes(x = tmax, fill = name)) + 
  geom_density(alpha = 0.3) 

```

What about boxplot?

```{r boxplot}
weather_df %>% 
  ggplot(aes(x = name, y = tmax)) +
  geom_boxplot()

```

Violin plot can be useful, sometimes. Can be used for data doublecheck.

```{r violin}
weather_df %>% 
  ggplot(aes(x = name, y = tmax)) +
  geom_violin()

```

Ridge plots? Useful for multiple group distribution.
Once extremely popular in 2017, only for 3 weeks.

```{r ridge plots}
weather_df %>% 
  ggplot(aes(x = tmax, y = name, fill = name)) +
  geom_density_ridges(alpha = 0.25)
#Careful about the x and y assignment
```

Let's try to save something. 
```{r save the plot}
weather_ridge = 
  weather_df %>% 
    ggplot(aes(x = tmax, y = name, fill = name)) +
    geom_density_ridges(alpha = 0.25)

ggsave("weather ridge plot.pdf", weather_ridge)
```

