---
title: "191001_visualization"
author: "Gavin Ko"
date: "10/1/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(viridis)
library(patchwork)
```

## Create the weather data
```{r weather data loading}
weather_df = 
  rnoaa::meteo_pull_monitors(c("USW00094728", "USC00519397", "USS0023B17S"), #three sites
                      var = c("PRCP", "TMIN", "TMAX"), #humidity, minimumtemp, highesttemp
                      date_min = "2017-01-01",
                      date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY", 
                      USC00519397 = "Waikiki_HA",
                      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
weather_df

```

## Plot a scatter
```{r adding lables and add specific data points, warning = FALSE}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .5) +
  labs(
    title = "Temp plot",
    xlab = "Minimum temperature (C)",
    ylab = "Maximum temperature",
    caption = "Data from NOAA via rnoaa pakcage"
  ) +
  scale_x_continuous(
    breaks = c(-15, 0, 20), #adding data points
    labels = c("-15C", "0C", "20C") #refine their respective lables
  ) + 
  scale_y_continuous( #you can also do data transformation in ggplot directly 
    trans = "sqrt"
  )
#Alternative of sqrt transforamtion : 
#scale_y_sqrt()

```

## Change the defult colors
Scale can be really useful in all aspects.
```{r adjusting colors, warning = FALSE}
ggplot_base = weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .5) +
  labs(
    title = "Temp plot",
    xlab = "Minimum temperature (C)",
    ylab = "Maximum temperature",
    caption = "Data from NOAA via rnoaa pakcage"
  ) +
  #scale_color_hue( #adjustment of colors
    #name = "Location", #legend main title
    #h = c(50,255) #a BAD way to put colors inside the points
  viridis::scale_color_viridis(
  name = "Location",
  labels = c("CentralPark","Waikiki","Waterhole"),
  discrete = T
  )


```
Since the original color of ggplot is barely acceptable, we now utilize viridis packages.

Let's do more about the legends parameters.

```{r change the legends}
ggplot_base + 
  theme_minimal() + 
  theme(legend.position = "bottom")
#Order does matter. Overarching things go first, then the tweeking stuff.

```

Inclass question: how do we check the NA values quickly?
```{r check NA data}
weather_df %>% 
  filter(is.na(tmin))

```

## more than one dataset

```{r stacking plots}
central_park = 
  weather_df %>% 
  filter(name == "CentralPark_NY")

waikiki = 
  weather_df %>% 
  filter(name == "Waikiki_HA")

waterhole =
  weather_df %>% 
  filter(name == "Waterhole_WA")

ggplot(data = waikiki, aes(x = date, y = tmax, color = name)) + 
  geom_point(aes(size = prcp)) + 
  geom_line(data = central_park)

```
(Brief aside about colors)

```{r specify the color}
waikiki %>% 
  ggplot(aes(x = date, y = tmax)) + 
  geom_point(color = "navy")
#color statement should be in point argument. Don't ask why, it's ggplot.
  
```
## Panels

```{r}
tmax_tmin_p = 
  weather_df %>% 
  ggplot(aes(x = tmax, y = tmin, color = name)) + 
  geom_point(alpha = .5) +
  theme(legend.position = "none")

prcp_dens_p = 
  weather_df %>% 
  filter(prcp > 0) %>% 
  ggplot(aes(x = prcp, fill = name)) + 
  geom_density(alpha = .5) + 
  theme(legend.position = "none")

tmax_date_p = 
  weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) + 
  geom_point(alpha = .5) +
  geom_smooth(se = FALSE) + 
  theme(legend.position = "bottom")

(tmax_tmin_p + prcp_dens_p) / tmax_date_p
```

## Data manipulation
```{r}
weather_df %>% 
  mutate(
    name = factor(name),
    name = fct_reorder(name, tmax) #reorder the factors
  ) %>% 
  ggplot(aes(x = name, y = tmax, color = name)) +
  geom_boxplot()

```

## Restructure then plot
```{r}
weather_df %>% 
  pivot_longer(
    tmax:tmin,
    names_to = "observation",
    values_to = "temp"
  ) %>% 
  ggplot(aes(x = temp, fill = observation)) + 
  geom_density(alpha = .5) +
  facet_grid(~name) + #create multiple comparison
  viridis::scale_fill_viridis(discrete = TRUE)

```





